Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)

### 2018-02-27 paper.tex
update

We use Scala language to demonstrate the idea and each step of transformations.
The reader should have moderate familiarity to Scala's syntax, such as
case classes, pattern matching and for comprehension.

@@

\begin{lstlisting}
def drive_step(todo: List[State], seen: Set[State]): Set[State] = todo match {
  case Nil => seen
  case hd::tl if seen.contains(hd) => drive_step(tl, seen)
  case hd::tl =>
    val new_time = tick(hd)
    hd match {
      case State(Let(x, App(f, ae), e), env, bstore, konts, time)
        if isAtomic(ae) =>
        val closures = atomicEval(f, env, bstore).toList
        val new_states = for (Clos(Lam(v, body), env_c) <- closures) yield {
          val frame = Frame(x, e, env)
          val baddr = allocBind(v, new_time)
          val new_env = env_c + (v -> baddr)
          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
          State(body, new_env, new_store, frame::konts, new_time)
        }
        drive_step(new_states ++ tl, seen + hd)
      case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
        continue(konts, atomicEval(ae, env, bstore), bstore, new_time, tl, seen + hd)
    }
}

def continue(konts: List[Frame], ds: Set[Storable], bstore: BStore, time: Time, 
             todo: List[State], seen: Set[State]): Set[State] = 
  konts match {
    case Nil => drive_step(todo, seen)
    case Frame(x, e, f_env)::konts =>
      val baddr = allocBind(x, time)
      val new_env = f_env + (x -> baddr)
      val new_store = bstore.update(baddr, ds)
      val new_state = State(e, new_env, new_store, konts, time)
      drive_step(new_state::todo, seen)
  }
\end{lstlisting}

@@

>>>>

We use Scala language to demonstrate the idea and each step of transformations.
We expect that readers have moderate familiarity to Scala's syntax, such as
case classes, pattern matching and for comprehension.

@@

\begin{lstlisting}
def aval(state: State, todo: List[State], seen: Set[State]): Set[State] = { 
  val newTime = tick(state)
  state match {
    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
      val newStates = for (Clos(Lam(v, body), c_env) <- aeval(f, env, bstore).toList) yield {
        val frame = Frame(x, e, env)
        val baddr = allocBind(v, newTime)
        val newEnv = c_env + (v -> baddr)
        val newStore = bstore.update(baddr, aeval(ae, env, bstore))
        State(body, newEnv, newStore, frame::konts, newTime)
      }   
      collect(newStates ++ todo, seen + state)
    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
      continue(konts, aeval(ae, env, bstore), bstore, newTime, todo, seen + state)
  }   
}

def collect(todo: List[State], seen: Set[State]): Set[State] = { 
  todo match {
    case Nil => seen
    case hd::tl if seen.contains(hd) => collect(tl, seen)
    case hd::tl => aval(hd, tl, seen)
  }   
}

def continue(konts: List[Frame], ds: Set[Storable], bstore: BStore, time: Time,
              todo: List[State], seen: Set[State]): Set[State] = konts match {
  case Nil => collect(todo, seen)
  case Frame(x, e, f_env)::konts =>
    val baddr = allocBind(x, time)
    val newEnv = f_env + (x -> baddr)
    val newStore = bstore.update(baddr, ds) 
    val newState = State(e, newEnv, newStore, konts, time)
    collect(newState::todo, seen)
}

def analyze(e: Expr): Set[State] = aval(inject(e), List(), Set()) 
\end{lstlisting}

@@

\begin{lstlisting}
case class Config(e: Expr, env: Env, store: BStore, time: Time)
case class ValStore(vs: Set[Storable], store: BStore, time: Time)
type Cache = Map[Config, ValStore]

type Continue = Function4[Set[Storable], BStore, Time, Cache, ValStore]

def tick(s: Config): Time = (s.e::s.time).take(k)

def aval(e: Expr, env: Env, bstore: BStore, time: Time, cache: Cache, continue: Continue): ValStore = {
  val config = Config(e, env, bstore, time)
  if (cache.contains(config)) {
    val ValStore(vs, bstore, time) = cache(config)
    return continue(vs, bstore, time, cache)
  }
  val newTime = tick(config)
  e match {
    case Let(x, App(f, ae), e) if isAtomic(ae) =>
      val closures = aeval(f, env, bstore).toList

      def nd(closures: List[Storable], accds: Set[Storable], accbstore: BStore, acctime: Time, cache: Cache): ValStore  = {
        closures match {
          case Nil =>
            val baddr = allocBind(x, acctime)
            val newEnv = env + (x -> baddr)
            val newStore = accbstore.update(baddr, accds)
            aval(e, newEnv, newStore, acctime, cache, (ds: Set[Storable], accbstore: BStore, acctime: Time, accCache: Cache) => {
                val newCache = accCache + (Config(e, newEnv, newStore, acctime) -> ValStore(ds, accbstore, acctime))
                continue(ds, accbstore, acctime, newCache)
            })
          case Clos(Lam(v, body), env_c)::rest =>
            val baddr = allocBind(v, newTime)
            val newEnv = env_c + (v -> baddr)
            val newStore = accbstore.update(baddr, aeval(ae, env, bstore)) //TODO: which store?
            aval(body, newEnv, newStore, newTime, cache, (ds: Set[Storable], accbstore: BStore, acctime: Time, accCache: Cache) => {
                val newCache = accCache + (Config(body, newEnv, newStore, newTime) -> ValStore(ds, accbstore, acctime))
                nd(rest, ds++accds, accbstore, acctime, newCache)
            })
        }
      }

      nd(closures, Set(), bstore, newTime, cache)
    case ae if isAtomic(ae) =>
      val ds = aeval(ae, env, bstore)
      continue(ds, bstore, newTime, cache)
  }
}

def done: Continue = (ds: Set[Storable], bstore: BStore, time: Time, cache: Cache) => {
  ValStore(ds, bstore, time)
}
def analyze(e: Expr): ValStore = aval(e, Map(), Store[BAddr, Storable](Map()), List(), Map[Config, ValStore](), done)
\end{lstlisting}
